---
title: "Lab 06: Join those tables!"
author: "Shafayet Shariar Hossain"
date: "3/1/2022"
output: html_document
---

## Load packages and data

```{r setup, message=FALSE}

library(tidyverse)

```


## Exercises 1

### Exercise 1.1

```{r dataload, message=FALSE}
songs <- read_csv("https://raw.githubusercontent.com/bwiernik/progdata-class/master/data/singer/songs.csv")
locations <- read_csv("https://raw.githubusercontent.com/bwiernik/progdata-class/master/data/singer/loc.csv")

(time <- as_tibble(songs) |> 
   rename(song = title))

(album <- as_tibble(locations) |> 
   select(title, everything()) |> 
   rename(album = release,
          song  = title))
```

 The songs for those we know its corresponding album
 
```{r}

time |> 
  inner_join(album, by = 'song')
  
```
 
 
### Exercise 1.2
 
 
Adding the corresponding albums to the time tibble, even if album info is not readily available. 
 
 
```{r}

time |> 
  left_join(album, by = 'song')
  
```


### Exercise 1.3


The songs for those we do have “year”, but not album info.


```{r}

time |> 
  anti_join(album, by = "song")


```


### Exercise 1.4

Artists that are in time but not in album.

```{r}

time |> 
  anti_join(album, by = "artist_name")

```


### Exercise 1.5

To make all the info available in one tibble.



```{r}

time |>
  full_join(album, by = 'song')


```


## Exercise 2


### Exercise 2.1

Three tibbles of data on the Lord of the Rings:


```{r data_load, message=FALSE}

fell <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Fellowship_Of_The_Ring.csv")
ttow <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Two_Towers.csv")
retk <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Return_Of_The_King.csv")


```


Stacking into a single tibble.


```{r}
bind_rows(fell, ttow, retk)
```


### Exercise 2.2

Races that are present in “The Fellowship of the Ring” (fell), but not in any of the other ones.

```{r}

fell |> 
  anti_join(ttow, by = "Race") |> 
  anti_join(retk, by = "Race")

```


## Exercise 3


### Exercise 3.1


```{r}

(y <-  tibble(x1 = LETTERS[1:3], x2 = 1:3))
(z <- tibble(x1 = c("B", "C", "D"), x2 = 2:4))

```

Rows that appear in both y and z


```{r}

intersect(y, z)

```

### Exercise 3.2


A data set to reflect that the data in y collected on Day 1, and z in Day 2


```{r}

union(
  mutate(y, day = "Day 1"),
  mutate(z, day = "Day 2")
)

```


### Exercise 3.3

Remove z rows from y.

```{r}
setdiff(y, z)

```



