---
title: "Lab 06: Join those tables!"
author: "Shafayet Shariar Hossain"
date: "3/1/2022"
output: html_document
---

## Load packages and data

```{r setup, message=FALSE}

library(tidyverse)

```


## Exercises 1

### Exercise 1.1

```{r}
songs <- read_csv("https://raw.githubusercontent.com/bwiernik/progdata-class/master/data/singer/songs.csv")
locations <- read_csv("https://raw.githubusercontent.com/bwiernik/progdata-class/master/data/singer/loc.csv")

(time <- as_tibble(songs) |> 
   rename(song = title))

(album <- as_tibble(locations) |> 
   select(title, everything()) |> 
   rename(album = release,
          song  = title))
```

 The songs for those we know its corresponding album
 
```{r}

time |> 
  inner_join(album, by = 'song')
  
```
 
 
### Exercise 1.2
 
 
Adding the corresponding albums to the time tibble, even if album info is not readily available. 
 
 
```{r}

time |> 
  left_join(album, by = 'song')
  
```


### Exercise 1.3


The songs for those we do have “year”, but not album info.


```{r}

time |> 
  anti_join(album, by = "song")


```


### Exercise 1.4

Artists that are in time but not in album.

```{r}

time |> 
  anti_join(album, by = "artist_name")

```


### Exercise 1.5

To make all the info available in one tibble.



```{r}

time |>
  full_join(album, by = 'song')


```


## Exercise 2


### Exercise 2.1

Three tibbles of data on the Lord of the Rings:


```{r}

fell <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Fellowship_Of_The_Ring.csv")
ttow <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Two_Towers.csv")
retk <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Return_Of_The_King.csv")


```


Stacking into a single tibble.


```{r}

```

