---
title: "Lab 03 Explore gapminder with ggplot2 and dplyr"
author: "Shafayet Shariar Hossain"
date: "02/08/2022"
output: html_document
---

## Load packages and data

```{r setup, message=FALSE}
library(tidyverse)
library(gapminder)
```


## Exercises

### Exercise 1.1


```{r filter gapminder to 3 countries in the 1970s}

filter(gapminder, country %in% c('Bangladesh','France','United States') & year %in% c(1970:1979))

```

### Exercise 1.2

```{r use the pipe operator |> to first do the above filter and then select the “country” and “gdpPercap” variables.}


gapminder |> 
  filter(country %in% c('Bangladesh','France','United States')) |> 
  filter(year %in% c(1970:1979)) |> 
  select(country, gdpPercap)

```

### Exercise 1.3

```{r a new variable in gapminder for the change in life expectancy from the previous measurement for that country}


new_gapminder <- gapminder |> 
  arrange(year) |>
  group_by(country) |> 
  mutate(lifeExp_change = lifeExp - lag(lifeExp)) |> 
  filter(lifeExp_change < 0)

new_gapminder

```

### Exercise 1.4

```{r max GDP per capita experienced by each country}

gapminder |> 
  group_by(country) |> 
  filter(gdpPercap == max(gdpPercap))


```

### Exercise 1.5

```{r a scatterplot of Canada’s life expectancy vs. GDP per capita}


ggplot(gapminder |> 
  filter(country %in% 'Canada')) +
  aes(x = gdpPercap, y = lifeExp) +
  geom_point()+
  scale_x_continuous(trans = 'log10', labels = scales::dollar_format())


```




### Exercise 2.1


```{r a summary table of descriptive statistics for two quantitative variables}

# I have chosen the gapminder data set. And the chosen variables are
# lifeExp & population


gapminder |>
  summarize(
    across(c(lifeExp,pop),
           list(
             mean = \(x) mean(x),
             median = \(x) median(x),
             sd = \(x) sd(x)
           )) 
  )
```

```{r a scatterplot of lifeExp and pop variables using ggplot()}

ggplot(gapminder |>
         select(lifeExp,pop)) +
  aes(x = pop, y = lifeExp)+
  geom_point()+
  scale_x_continuous(trans = 'log10', labels = scales::comma_format())



```

### Exercise 2.2


```{r a summary table giving the sample size and descriptive statistics for the quantitative variable by group.}

# I have chosen the gapminder data set. And the chosen variables are
# lifeExp & country

gapminder |>
  group_by(country) |>
  summarize(count = n(),
            lifeExp_mean = mean(lifeExp),
            lifeExp_median = median(lifeExp),
            lifeExp_sd = sd(lifeExp)
  )

```


```{r Raincloud Plot}

ggplot(gapminder |>
         filter(country %in% c('United States','China','Germany'))) +
  aes(x = lifeExp, y = country) + 
  geom_jitter(height = 0.2, width = 0.5) +
  geom_boxplot(alpha = 0.3, width = 0.3) +
  ggdist::stat_slab(position = position_nudge(y = 0.3),
                    alpha = 0.5, size = 0.1, height = 0.5) 
  

```

### Bonus



```{r}

filter(gapminder, country == c("Rwanda", "Afghanistan"))

```

From the tibble above we find 12 rows of data. However, we know for a fact that our tibble should have 24 rows of data (12 for each country). Therefore, the code above didn't work. The reason being that when R is evaluating the relation country == c("Rwanda", "Afghanistan") it is evaluating 1st element of country with the 1st element of our given vector. In this case Rwanda and for that reason we do not have the 1st actual row of Afghanistan. The second value of the country vector is Afghanistan which is matched with the second element of our given vector Afghanistan. Therefore, we have that in our tibble. Similarly, the rest of the elements in the country vector gets matched with our given vector. This is how we lost half the elements of each country. 

The correct way to do would be the following

```{r}
filter(gapminder, country %in% c("Rwanda", "Afghanistan"))

```

