---
title: "Lab 02 - Plastic waste"
author: "Shafayet Shariar Hossain"
date: "02/01/2022"
output: html_document
---

## Load packages and data

```{r setup, message=FALSE}
library(tidyverse)
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv(here::here("data/plastic-waste.csv"))
```

## Exercises

### Exercise 1.1

Remove this text, and add your answer for Exercise 1 here.

```{r plastic-waste-continent}

ggplot(data = plastic_waste) +
  aes(x = plastic_waste_per_cap) +
  geom_histogram(binwidth = 0.2, na.rm = TRUE) +
  facet_wrap(~continent, nrow = 2)


ggplot(data = plastic_waste) +
  aes(x = plastic_waste_per_cap) + 
  geom_density(na.rm = TRUE)+
  facet_wrap(~ continent, nrow = 2)

ggplot(data = plastic_waste) +
  aes(x = plastic_waste_per_cap) + 
  geom_histogram(na.rm = TRUE, binwidth = 0.2, aes(y = ..density..)) + 
  geom_density(na.rm = TRUE) + 
  facet_wrap(~ continent, nrow = 2)
```

### Exercise 1.2

```{r plastic-waste-density}

ggplot(data = plastic_waste) +
  aes(x = plastic_waste_per_cap, color = continent) + 
  geom_density(na.rm = TRUE)+
  facet_wrap(~ continent, nrow = 2)


ggplot(data = plastic_waste) +
  aes(x = plastic_waste_per_cap, color = continent, fill = continent) + 
  geom_density(na.rm = TRUE, alpha = 0.5)+
  facet_wrap(~ continent, nrow = 2)


```

### Exercise 1.3

Remove this text, and add your answer for Exercise 3 here.



### Exercise 2.1

Remove this text, and add your answer for Exercise 4 here.

```{r plastic-waste-violin}

ggplot(data = plastic_waste) +
  aes(y = plastic_waste_per_cap, x = continent, fill = continent) + 
  geom_boxplot(na.rm = TRUE, alpha = 0.5)
  
```

### Exercise 2.2

Remove this text, and add your answer for Exercise 5 here.

```{r plastic-waste-mismanaged}

ggplot(data = plastic_waste) +
  aes(y = plastic_waste_per_cap, x = continent, fill = continent, color = continent) + 
  geom_jitter(height = 0.2, width = 0.2, na.rm = TRUE) +
  geom_boxplot(na.rm = TRUE, alpha = 0.5) +
  ylim(0,1)

```

### Exercise 2.3

Remove this text, and add your answer for Exercise 6 here.

```{r plastic-waste-mismanaged-continent}

ggplot(data = plastic_waste) +
  aes(x = plastic_waste_per_cap, y = continent, fill = continent, color = continent) + 
  geom_jitter(height = 0.2, width = 0.2, na.rm = TRUE) +
  geom_boxplot(na.rm = TRUE, alpha = 0.5, width = 0.5, color = 'black') +
  ggdist::stat_slab(na.rm = TRUE, position = position_nudge(y = 0.2),
                    alpha = 0.5, size = 0.1, height = 0.5) +
  xlim(0,1)

```

### Exercise 2.4

Remove this text, and add your answer for Exercise 7 here.


### Exercise 3.1

```{r relationship between plastic waste per capita and mismanaged plastic waste per capita}

ggplot(data = plastic_waste) +
  aes(x = mismanaged_plastic_waste_per_cap, y = plastic_waste_per_cap) + 
  geom_point(na.rm = TRUE) +
  ylim(0,1)

```

### Exercise 3.2

```{r Color the points in the scatterplot by continent}


ggplot(data = plastic_waste) +
  aes(x = mismanaged_plastic_waste_per_cap, y = plastic_waste_per_cap, color = continent) + 
  geom_point(na.rm = TRUE) +
  ylim(0,1)
```


### Exercise 3.3

```{r Relationship between plastic waste per capita and total population}

ggplot(data = plastic_waste) +
  aes(x = total_pop, y = plastic_waste_per_cap) +
  geom_point(na.rm = TRUE) +
  scale_x_continuous(trans = 'log10', labels = scales::comma_format()) +
  ylim(0,1)+
  geom_smooth(na.rm = TRUE)


```


```{r Relationship between plastic waste per capita and coastal population}

ggplot(data = plastic_waste) +
  aes(x = coastal_pop, y = plastic_waste_per_cap) +
  geom_point(na.rm = TRUE) +
  scale_x_continuous(trans = 'log10', labels = scales::comma_format()) +
  ylim(0,1)+
  geom_smooth(na.rm = TRUE)

```



### Exercise 4

Remove this text, and add your answer for Exercise 8 here.

```{r recreate-viz}
new_plastic_waste <- plastic_waste %>% 
  drop_na() %>% 
  mutate(Coastal_pop_proportion = coastal_pop/total_pop)

ggplot(new_plastic_waste) +
  aes(x = Coastal_pop_proportion, y = plastic_waste_per_cap) +
  geom_point(na.rm = TRUE, size = 1.5, aes(color = continent)) +
  geom_smooth(na.rm = TRUE, color = 'black') +
  ylim(0,0.75) +
  xlab('Coastal population proportion (Coastal / total population)') +
  ylab('Plastic waste per capita') +
  labs(title = 'Plastic waste vs. coastal population proportion',
       subtitle = 'by continent') +
  scale_color_discrete('Continent')
  

```

